MISUSES AND INACCURACIES OF PAGEL'S LAMBDA ML ESTIMATION METHODS

### 2020-04-16 Manuscript/Manuscript-1-DCA.tex
vsn 1-dca

Investigating macroevolutionary patterns of trait variation requires a
phylogenetic perspective, because the shared ancestry among species
generates statistical non-independence (Felsenstein 1985; Harvey and
Pagel 1991). Accounting for this evolutionary non-independence is the
purview of \emph{phylogenetic comparative methods} (PCMs); a suite of
analytical tools that condition the data on the phylogeny through the
course of statisical evaluations of phenotypic trends (e.g., Grafen
1989; Garland and Ives 2000; Rohlf 2001; Butler and King 2004). The past
several decades have witnessed a rapid expansion in the development of
PCMs to address an ever-growing set of macroevolutionary hypotheses
(Martins and Hansen 1997; O'Meara et al. 2006; Revell and Harmon 2008;
Beaulieu et al. 2012; Adams 2014b,a; Adams and Collyer 2018). These
methods are predicated on the notion that phylogenetic signal -- the
tendancy for closely related species to display similar trait values --
is present in cross-species datasets (Abouheif 1999; Pagel 1999;
Blomberg et al. 2003; Munkemuller et al. 2012). Indeed, under numerous
evolutionary models, phylogenetic signal is to be expected, as
stochastic character change along the hierarchical structure of the tree
of life generates trait covaration among related taxa (see Felsenstein
1985; Blomberg et al. 2003; Revell et al. 2008). \hfill\break

In many macroevolutionary studies, it is often of interest to quantify
the degree to which phylogenetic signal is displayed in continuous
traits. Several analytical tools have been developed for this purpose
(e.g., Gittleman and Kot 1990; Abouheif 1999; Pagel 1999; Blomberg et
al. 2003; Klingenberg and Gidaszewski 2010; Adams 2014a), which differ
primarily in how they characterize the phylogenetic dependency of trait
variation among taxa. One commonly used statistical measure,
\emph{Kappa} (\(K\)), expresses the strength of phylogenetic signal as
the ratio of observed trait variation to the trait variation conditioned
on the phylogeny; scaled by what is expected under Brownian motion given
the phylogeny's size and shape (Blomberg et al. 2003; also Adams 2014a).
Another approach, Pagel's \(\lambda\) (Pagel 1999), uses maximum
likelihood to fit the data to the phylogeny under some model of
evolutionary change (typically Brownian motion). The inclusion of a
scaling parameter, \(\lambda\), transforms the lengths of the internal
branches of the phylogeny to improve the fit, and this parameter
describes the degree of phylogenetic signal in the dataset (Pagel 1999;
Freckleton et al. 2002). Pagel's \(\lambda\) also has the advantage that
it may be included when estimating the association of traits in a
phylogenetic context, meaning that one may account for the degree of
phylogenetic signal while conducting phylogenetic regression or ANOVA
(see Freckleton et al. 2002). \hfill\break

Several studies have investigated the statistical properties of methods
for estimating phylogenetic signal under various conditions (e.g.,
Munkemuller et al. 2012; Pavoine and Ricotta 2012; Diniz-Filho et al.
2012; Molina-Venegas and Rodriguez 2017; see also Revell et al. 2008;
Revell 2010). These have largely focused on the ability of methods to
detect the presence of phylogenetic signal (i.e., type I and type II
error rates) under complex models of evolutionary change, across a range
of phylogeny sizes, and with varying degrees of phylogenetic uncertainty
or unresolved topologies. In terms of parameter estimation, Revell
(2010) confirmed that regression parameters were accurately estimated
when \(\lambda\) was included during phylogenetic regression, and
Munkmuller et al.~(2012) found that estimates of phylogenetic signal
obtained using various measures generally increased when input levels of
phylogenetic signal were stronger. However, the precision of those
estimates could not be determined, because the input levels of
phylogenetic signal were simulated via a scaling factor (\(w\)) that was
not directly comparable to the measures of phylogenetic signal being
compared (see Munkemuller et al. 2012). One study (Boettiger et al.
2012) found that estimates of Pagel's \(\lambda\) displayed less
variation when data were simulated on a large phylogeny (\(N=281\)) as
compared to a small one (\(N=13\)), concluding that insufficient data
(i.e., the number of species) was the underlying cause of the lack of
precision in parameter estimation. However, this conclusion assumes that
parameter estimation remains equally precise across its range of values
(\(\lambda = 0 \to 1\)); an assumption that to date, has not been
investigated. Thus, despite widespread use of Pagel's (1999) \(\lambda\)
in macroevolutionary studies, at present, we still lack a general
understanding of the precision with which \(\lambda\) can estimate
levels of phylogenetic signal in phenotypic datasets. \hfill\break

In this study, we evaluate the precision of Pagel's \(\lambda\) to
estimate known levels of phylogenetic signal in phenotypic data. First
we use computer simulations across differing numbers of species, and
differing input levels of phylogenetic signal to explore when
\(\lambda\) captures known levels of phylogenetic signal, and under what
circumstances. Additionally, we use simulations to determine how the
inclusion of \(\lambda\) in phylogenetic ANOVA and regression (i.e.,
PGLS) affects parameter estimation, and whether estimates of
phylogenetic signal in PGLS are accurate. We then survey the recent
macroevolutionary literature for published papers containing estimates
of \(\lambda\) from empirical datasets, and compare these empirical
estimates to patterns gleaned from our computer simulations. In general
we find that while PGLS parameters are accurately estimated with the
inclusion of phylogenetic signal, estimates of \(\lambda\) are not.
Thus, empirical studies should interpret the degree of phylogenetic
signal in phenotypic data with caution when using this parameter.
\textbf{something more? Let's wait for final data to decide.}

\hypertarget{methods-and-results}{%
\section{Methods and Results}\label{methods-and-results}}

\hypertarget{simulated-trait}{%
\subsection{\texorpdfstring{\emph{Simulated
trait}}{Simulated trait}}\label{simulated-trait}}

To assess the accuracy of Pagel's lambda estimations, we simulated
pure-birth phylogenies of variable size, ranging in tip number from 32
to 1024. We then scaled the simulated phylogenies by lambda values
ranging from 0 to 1 (0.05 intervals; 50 trees per lambda value per tree
size) with which we generated trait data with known lambdas by
simulating a continuous variable on each scaled phylogeny under Brownian
motion. We then estimated lambda values from these data using
phylogenetic generalized least squares (PGLS) to compare against the
known lambda values. \hfill\break

To visualize the accuracy of the lambda estimation methods, we first
plotted known lambdas (input lambdas) against the estimated lambdas
(Figure 1). This plot demonstrates the rampant inaccuracy of estimating
lambdas on phylogenies with fewer than 200 tips, as the spread of data
in the upper panels in Figure 1 is remarkably wide. We also see that the
widest spread of estimated lambdas is observed near the center of each
plot corresponding to intermediate values of known lambda. Lastly, we
see a slight tendency for the PGLS estimation methods to underestimate
lambda, especially for known lambda values below 0.5 as can be seen by
the numerous data points along the x axis for the smaller phylogenies
analyzed (n tips \textless{} 200). \hfill\break

{[}insert Figure 1 here{]}

\hypertarget{simulated-anova-and-regressions}{%
\subsection{\texorpdfstring{\emph{Simulated ANOVA and
Regressions}}{Simulated ANOVA and Regressions}}\label{simulated-anova-and-regressions}}

To measure the statistical performance of PGLS lambda estimation methods
when applied to ANOVA and regression analyses, we used the above
generated data (independent variables) to simulate a second set of trait
variables (depedendent variables) across the range of correlation
strengths (betas ranging from 0 to 1 at intervales of 0.25). We then
used PGLS to estimate phylogenetic signal of the dependent variable and
the slope coefficient for the regression between the dependent and
independent variables. We also calculated F- and p-values from the
regression analyses. Finally, we again fit the dependent variable to the
independent variable while holding the lambda value at 0 to approximate
an ordinary least squares (OLS) apprach. \hfill\break

We compared the estimated slope coefficients across variable input beta
values to assess the ability of PGLS to identify significant
correlations (Figure 2). All distribution groups center around the 1:1
relationship between input beta values and estimated slopes. However the
variance around that mean is substantial, with estimates ranging from
approximately -2 to 4 in datasets with 32 taxa and a known relationship
of (beta = 1). This variance of the estimated slopes only becomes
reasonable when phylogenies have over 200 tips, similarly to what we saw
in the first simulation analysis in this study. \hfill\break

@@

Surprisingly, the estimated lambda values of the dependent variable do
not correspond with input lambda values that characterize the
independent variable (Figure S1). This disconnect then resulted in no
appreciable difference of slope estimates across input lambda values
(Figure S2), nor do we see substantial differences in slope estimates
between the PGLS and OLS analyses (results not shown). Thus, we show
that phylogenetic signal present in one variable does not translate to
phylogenetic signal in a second, even highly correlated, dependent
variable. \hfill\break

Scaling and trait generation procedures for all simulation methods were
repeated with symmetrical and ladder phylogenies of variable size.
Results generated using these variable phylogeny shapes were consistent
with the pure-birth phylogeny results presented above and can be found
in the Supplemental Materials. All analyses were performed in R v3.6.0
(R Core Team 2019) using the packages \texttt{geiger} (Harmon et al.
2008) and \texttt{caper} (Orme et al. 2013), and the corresponding
scripts can be found in the Supplemental Materials.

\hypertarget{meta-analysis-of-empirical-results}{%
\subsection{\texorpdfstring{\emph{Meta-Analysis of Empirical
Results}}{Meta-Analysis of Empirical Results}}\label{meta-analysis-of-empirical-results}}

To understand the extent of this problematic estimation method in
application, we performed a meta-analysis of studies published in 2019
that cited Pagel's 1997 manuscript (Pagel 1999). The list of manuscripts
was compiled through Google Scholar on Jan 23, 2020 and totaled 341
manuscripts. For each study, we extracted any published lambda
estimates, along with the size of the phylogeny used in the analysis. We
also noted whether authors reported confidence intervals, significance
tests assessing difference of the lambda estimate from 0 or 1, and
whether authors interpreted biological meaning from the magnitude of the
estimated lambda. For studies that reported more than one lambda
estimate, we also noted if the authors compared the lambdas against one
another, and whether that was accompanied with an appropriate
statistical test between the estimated lambda values. \hfill\break

@@

@@

{[}This part is obviously not written yet{]} \hfill\break

General conclusions :Using the estimated lambda values from pgls are not
useful. The questions of whether or not signal exists is appropriate,
but inferring more from lambda \emph{magnitude} is inappropriate.
\hfill\break

@@

@@

@@

\leavevmode\hypertarget{ref-DinizFilho2012}{}%
Diniz-Filho, J. A. F., T. Santos, T. F. Rangel, and L. M. Bini. 2012. A
comparison of metrics for estimating phylogenetic signal under
alternative evolutionary models. Genetics and molecular biology
35:673--679. SciELO Brasil.

@@

\leavevmode\hypertarget{ref-Gittleman1990}{}%
Gittleman, J. L., and M. Kot. 1990. Adaptation: Statistics and a null
model for estimating phylogenetic effects. Systematic Zoology
39:227--241. Society of Systematic Zoology.

@@

\leavevmode\hypertarget{ref-Klingenberg2010}{}%
Klingenberg, C. P., and N. A. Gidaszewski. 2010. Testing and quantifying
phylogenetic signals and homoplasy in morphometric data. Systematic
biology 59:245--261. Oxford University Press.

@@

\leavevmode\hypertarget{ref-MolinaVenegas2017}{}%
Molina-Venegas, R., and M. A. Rodriguez. 2017. Revisiting phylogenetic
signal; strong or negligible impacts of polytomies and branch length
information? BMC evolutionary biology 17:53. BioMed Central.

@@

\leavevmode\hypertarget{ref-Pavoine2012}{}%
Pavoine, S., and C. Ricotta. 2012. Testing for phylogenetic signal in
biological traits: The ubiquity of cross-product statistics. Evolution:
International Journal of Organic Evolution 67:828--840. Wiley Online
Library.

@@

>>>>

Investigating macroevolutionary patterns of trait variation requires a
phylogenetic perspective, because the shared ancestry among species
generates statistical non-independence (Felsenstein 1985; Harvey and
Pagel 1991). Accounting for this evolutionary non-independence is the
purview of \emph{phylogenetic comparative methods} (PCMs); a suite of
analytical tools that condition the data on the phylogeny through the
course of statisical evaluations of phenotypic trends (e.g., Grafen
1989; Garland and Ives 2000; Rohlf 2001; Butler and King 2004). The past
several decades have witnessed a rapid expansion in the development of
PCMs to address an ever-growing set of macroevolutionary hypotheses
(Martins and Hansen 1997; O'Meara et al. 2006; Revell and Harmon 2008;
Beaulieu et al. 2012; Adams 2014b,a; Adams and Collyer 2018). These
methods are predicated on the notion that phylogenetic signal -- the
tendancy for closely related species to display similar trait values --
is present in cross-species datasets (Felsenstein 1985; Pagel 1999;
Blomberg et al. 2003). Indeed, under numerous evolutionary models,
phylogenetic signal is to be expected, as stochastic character change
along the hierarchical structure of the tree of life generates trait
covaration among related taxa (see Felsenstein 1985; Blomberg et al.
2003; Revell et al. 2008). \hfill\break

For many macroevolutionary analyses, it is often of interest to quantify
the degree to which phylogenetic signal is displayed in continuous
traits. Several analytical tools have been developed for this purpose
(e.g., Gittleman and Kot 1990; Abouheif 1999; Pagel 1999; Blomberg et
al. 2003; Klingenberg and Gidaszewski 2010; Adams 2014a), which differ
primarily in how they characterize the phylogenetic dependency of trait
variation among taxa. One commonly used statistical measure,
\emph{Kappa} (\(K\)), expresses the strength of phylogenetic signal as
the ratio of observed trait variation to the trait variation conditioned
on the phylogeny; scaled by what is expected under Brownian motion given
the phylogeny's size and shape (Blomberg et al. 2003; also Adams 2014a).
Another approach, Pagel's \(\lambda\) (Pagel 1999), uses maximum
likelihood to fit the data to the phylogeny under some model of
evolutionary change (typically Brownian motion). The inclusion of a
scaling parameter, \(\lambda\), transforms the lengths of the internal
branches of the phylogeny to improve the fit, and this parameter
describes the degree of phylogenetic signal in the dataset (Pagel 1999;
Freckleton et al. 2002). Pagel's \(\lambda\) also has the advantage that
it may be included when estimating the association of traits in a
phylogenetic context, meaning that one may account for the degree of
phylogenetic signal while conducting phylogenetic regression or ANOVA
(see Freckleton et al. 2002). \hfill\break

Several studies have investigated the statistical properties of methods
for estimating phylogenetic signal under various conditions (e.g.,
Munkemuller et al. 2012; Pavoine and Ricotta 2012; Diniz-Filho et al.
2012; Molina-Venegas and Rodriguez 2017; see also Revell et al. 2008;
Revell 2010). These have largely focused on the ability of methods to
detect the presence of phylogenetic signal (i.e., type I and type II
error rates) under complex models of evolutionary change, across a range
of phylogeny sizes, and with varying degrees of phylogenetic uncertainty
or unresolved topologies. In terms of parameter estimation, Revell
(2010) found that regression parameters were accurately estimated when
\(\lambda\) was included during phylogenetic regression, and Munkmuller
et al.~(2012) demonstrated that estimates of phylogenetic signal
obtained using various measures generally increased when input levels of
phylogenetic signal were stronger. However, the precision of those
estimates could not be determined, because the input levels of
phylogenetic signal were simulated via a scaling factor (\(w\)) that was
not directly comparable to the measures of phylogenetic signal being
compared (see Munkemuller et al. 2012). One study (Boettiger et al.
2012) found that estimates of Pagel's \(\lambda\) displayed less
variation when data were simulated on a large phylogeny (\(N=281\)) as
compared to a small one (\(N=13\)), concluding that insufficient data
(i.e., the number of species) was the underlying cause of the lack of
precision in parameter estimation. However, this conclusion assumes that
the precision of parameter estimation remains constant across the range
of values (\(\lambda = 0 \to 1\)); an assumption that to date, has not
been verified. Thus, despite widespread use of Pagel's (1999)
\(\lambda\) in macroevolutionary studies, at present, we still lack a
general understanding of the precision with which \(\lambda\) can
estimate levels of phylogenetic signal in phenotypic datasets.
\hfill\break

In this study, we evaluate the precision of Pagel's \(\lambda\) to
estimate known levels of phylogenetic signal in phenotypic data. First
we use computer simulations across differing numbers of species,
differently shaped phylogenies, and differing input levels of
phylogenetic signal, to explore the degree to which \(\lambda\)
correctly identifies known levels of phylogenetic signal, and under what
circumstances. For comparison, we also evaluate estimates of \(Kappa\)
and its empirically-derived effect size (\(Z_K\)) for this purpose.
Additionally, we use simulations to determine how the inclusion of
\(\lambda\) in phylogenetic regression and ANOVA (i.e., PGLS) affects
parameter estimation, and whether levels of phylogenetic signal
estimated in a PGLS framework are accurate. We then survey the recent
macroevolutionary literature for published papers containing estimates
of \(\lambda\) from empirical datasets, and compare these empirical
estimates to patterns gleaned from our computer simulations. In general
we find that while PGLS parameters (e.g., \(\beta\)) are accurately
estimated with the inclusion of phylogenetic signal, estimates of
\(\lambda\) are not. Further, we find that estimates of \(\lambda\) vary
widely for a given input value of phylogenetic signal, while effect
sizes (\(Z_K\)) obtained from \(Kappa\) display far less variation, and
increase in a steady (\textbf{uniform?}) manner with increasing
phylogenetic signal. We conclude that Pagel's \(\lambda\) should not be
used for characterizing and comparing the degree of phylogenetic signal
in phenotypic datasets, while comparisons of effect sizes obtained from
\(Kappa\) hold considerable promise in this regard.

\hypertarget{materials-and-methods}{%
\section{Materials and Methods}\label{materials-and-methods}}

\hypertarget{simulations}{%
\subsection{\texorpdfstring{\emph{Simulations}}{Simulations}}\label{simulations}}

We conducted a series of computer simulations to evaluate the ability of
Pagel's (1999) \(\lambda\) to estimate known levels of phylogenetic
signal. Our primary simulations were based on pure-birth phylogenies;
however, we also evaluated patterns on both balanced and pectinate trees
to determine the extent to which tree shape affected our findings (see
Supplemental Material). Our first set of simulations evaluated the
extent to which values of \(\lambda\) estimated from the data
corresponded with actual input levels of phylogenetic signal. For these
simulations we generated 50 pure-birth phylogenies at each of six
different tree sizes, ranging from 32 to 1024 taxa (\(n=2^5 - 2^{10}\)).
Next, we rescaled the simulated phylogenies by multiplying the internal
branches by \(\lambda\). We used a set of values that encompassed the
entire range of \(\lambda\): (\(\lambda_{in} = 0.0 - 1.0\); in 21
intervals of 0.05 units), resulting in 1050 scaled phylogenies at each
level of species richness (\(n\)). Continuous traits were then simulated
on each phylogeny under a Brownian motion model of evolution to obtain
datasets with known and differing levels of phylogenetic signal. These
varied from datasets with no phylogenetic signal (when
\(\lambda_{in} =0\)) to datsets with phylogenetic signal corresponding
to what was expected under Brownian motion (when \(\lambda_{in} =1\)).
\hfill\break

Using the simulated phylogenies we estimated the phylogenetic signal in
each dataset using Pagel's \(\lambda\). Next, for the 1050 datasets at
each level of species richness (\(n\)), we assessed the relationship
between \(\lambda_{in}\) and \(\lambda_{est}\) using regression.
Additionally, the precision of \(\lambda_{est}\) was approximated by
observing both the range and the variation of \(\lambda_{est}\) obtained
at each input level of phylogenetic signal (\(\lambda_{in}\)). Finally,
for comparison we estimated the degree of phylogenetic signal in each
dataset using \(Kappa\) (Blomberg et al. 2003; Adams 2014a), and
calculated its effect size (\(Z_K\): sensu Adams and Collyer 2016,
2019a) for each dataset. Here, the effect size was estimated as:
\(Z_K=\frac{K_{obs}-\mu_K}{\sigma_K}\), where \(K_{obs}\) was the
observed \(Kappa\), and \(\mu_K\) and \(\sigma_K\) were the mean and
standard deviation of the empirical sampling distribution of values
obtained from the permutation distribution (for discussion see: Adams
and Collyer 2019b). Because \(Kappa\) and \(Z_K\) differ in scale from
\(\lambda\), both measures were linearly normalized to a standard
uniform distribution (\(0\rightarrow1\)), and the range and variation in
the set of normalized \(Kappa\) and \(Z_K\) values at each input level
of phylogenetic signal (\(\lambda_{in}\)) were estimated for comparison.
\hfill\break

Our second set of simulations evaluated the extent to which values of
\(\lambda\) estimated in PGLS regression and ANOVA (i.e., \(Y\sim{X}\))
corresponded with actual input levels of phylogenetic signal in the
response variable. As before we generated 50 pure-birth phylogenies at
each of six levels of species richness (\(n=2^5 - 2^{10}\)), and
rescaled each with a set of input \(\lambda\) values
(\(\lambda_{in} = 0.0 - 1.0\); in 21 intervals of 0.05 units). Next, an
independent variable \(X\) was simulated on each phylogeny under a
Brownian motion model of evolution (for PGLS regression). For
phylogenetic ANOVA, random groups (\(X\)) were obtained by simulating a
discrete (binary) character on each phylogeny. Next, the dependent
variable was simulated in such a manner as to contain a known
relationship with \(X\) plus random error containing phylogenetic
signal. This was accomplished as: \(Y=\beta{X}+\epsilon\). Here, the
association between \(Y\) and \(X\) was modeled using a range of values
(\(\beta=(0,0.25, 0.5, 0.75,1)\)), and the residual error was modeled to
contain phylogenetic signal simulated under a Brownian motion model of
evolution (\(\epsilon=\mathcal{N}(\mu=0,\sigma=\mathbf{C})\): see Revell
2010 for a similar simulation design). For each dataset, the fit of the
phylogenetic regression was estimated using maximum likelihood, and
parameter estimates (\(\beta_{est}\) and \(\lambda_{est}\)) were
obtained and evaluated as above. All analyses were performed in R v3.6.0
(R Core Team 2019) using the packages \texttt{geiger} (Harmon et al.
2008), \texttt{caper} (Orme et al. 2013), \texttt{phytools} (Revell
2012), and \texttt{geomorph} (Adams and Otárola-Castillo 2013; Adams et
al. 2020). R-scripts are found in the Supplemental Material.

\hypertarget{literature-survey}{%
\subsection{\texorpdfstring{\emph{Literature
Survey}}{Literature Survey}}\label{literature-survey}}

To determine how Pagel's \(\lambda\) is commonly utilized in empirical
studies, we conducted a literature survey. From Google.scholar we
obtained a list of all papers published in 2019 that used \(\lambda\);
resulting in 341 studies. For each study, we extracted all
\(\lambda_{est}\), the size of the phylogeny (\(n\)), and noted whether
authors reported confidence intervals or performed significance tests
assessing difference of \(\lambda_{est}\) from either zero or one. We
also noted whether biological interpretations based on \(\lambda_{est}\)
were made, and for studies that reported more than one
\(\lambda_{est}\), we also noted whether these were compared in some
manner, and whether such comparisons were accompanied with statistical
tests between \(\lambda_{est}\). \hfill\break

\hypertarget{results}{%
\section{Results}\label{results}}

\hypertarget{simulations-1}{%
\subsection{\texorpdfstring{\emph{Simulations}}{Simulations}}\label{simulations-1}}

lambda: (Fig 1) not good predictor of input physignal (regression),
particularly at low sample sizes. Also, similar to early results of
Boettiger (but greatly expanded) range and variation of lambda very
large at low N, and generally decreases with larger N. However, the
precision was not uniform across the range of lambda-in. Variation in
lambda-est was greatest at intermediate levels of physig, and descreased
at lower and higher values. Thus, the overall dispersion across the
range of input values resembled an asymmetric lens (\textbf{NAME?}),
with less variation at lambda-in close to 1, greatest variation in
middle, and less (but substantial) variation near lambda-in of zero.
Notably, even at very large species richness (n=1024) the range of
lam-est was quite large: e.g., l-in 0.5 has range of 45\% of all
possible values. Thus, lambda-est is not very precise in terms of
estimating the true l-in parameter, meaning that using lambda to
biologically interpret the `strength' of phylogenetic signal in a
dataset, or compare between datasets, is fraught with difficulty.
\hfill\break

{[}insert Figure 1 here{]} \hfill\break

By contrast, (FIg 2) Kappa. Much smaller range and variation as compared
with lambda, but with larger l-in, variation in kappa does increase.
However, this is not the case with Z-K. Z-K increases predictably
(linearly, progressively\ldots) with L-in, and importantly variation in
Z-K remains nearly constant across the range of input values. this means
that Z-K is equally precise in estimating the strength of phylogenetic
signal, and may be more adept at facilitating useful statements on teh
relative strength of phylogenetic signal within and across datasets.
\hfill\break

\textbf{NOTE: Fig 2 is 4 panels for N=128: Lambda, Kappa, Z-K,
Z-k-normalized.}

@@

\hypertarget{literature-survey-1}{%
\subsection{\texorpdfstring{\emph{Literature
Survey}}{Literature Survey}}\label{literature-survey-1}}

@@

\hypertarget{results-1}{%
\section{Results}\label{results-1}}

@@

1: summary paragraph

2: expand on Lambda.. lambda innacurate, not precise, level of precision
varies with input physig (worse in mid-range). Patterns worse with PGLS,
though beta still estimated properly. Conclusion, lambda not overly
useful.

3: By contrast, effect size Z-K useful, equally precise across range of
values. Can be used to characterize the strength of physignal, and
because robust to input levels, etc. may be used to compare across
datasets.

Somewhere, recognize that this is somewhat `backwards' from prior
recommendations where Kappa had somewhat lower performance in terms of
type I and type II error (which?? I forget). However, recall that those
studies did not examine the precision of the estimates. Nor was Z-k
included, because it was not yet invented. So Use of Z-k should make
good sense here.

Closing paragraph.

@@

\leavevmode\hypertarget{ref-AdamsCollyer2019b}{}%
Adams, D. C., and M. L. Collyer. 2019a. Comparing the strength of
modular signal, and evaluating alternative modular hypotheses, using
covariance ratio effect sizes with morphometric data. Evolution
73:2352--2367.

\leavevmode\hypertarget{ref-AdamsCollyer2016}{}%
Adams, D. C., and M. L. Collyer. 2016. On the comparison of the strength
of morphological integration across morphometric datasets. Evolution
70:2623--2631.

@@

\leavevmode\hypertarget{ref-AdamsCollyer2019}{}%
Adams, D. C., and M. L. Collyer. 2019b. Phylogenetic comparative methods
and the evolution of multivariate phenotypes. Annual Review of Ecology,
Evolution, and Systematics 50:405--425.

\leavevmode\hypertarget{ref-AdamsGeomorph}{}%
Adams, D. C., M. L. Collyer, and A. Kaliontzopoulou. 2020. Geomorph:
Software for geometric morphometric analyses. R package version 3.2.1.

\leavevmode\hypertarget{ref-AdamsOtarola2013}{}%
Adams, D. C., and E. Otárola-Castillo. 2013. Geomorph: An r package for
the collection and analysis of geometric morphometric shape data.
Methods in Ecology and Evolution 4:393--399.

@@

\leavevmode\hypertarget{ref-DinizFilho2012}{}%
Diniz-Filho, J. A. F., T. Santos, T. F. Rangel, and L. M. Bini. 2012. A
comparison of metrics for estimating phylogenetic signal under
alternative evolutionary models. Genetics and Molecular Biology
35:673--679.

@@

\leavevmode\hypertarget{ref-Gittleman1990}{}%
Gittleman, J. L., and M. Kot. 1990. Adaptation: Statistics and a null
model for estimating phylogenetic effects. Systematic Zoology
39:227--241.

@@

\leavevmode\hypertarget{ref-Klingenberg2010}{}%
Klingenberg, C. P., and N. A. Gidaszewski. 2010. Testing and quantifying
phylogenetic signals and homoplasy in morphometric data. Systematic
biology 59:245--261.

@@

\leavevmode\hypertarget{ref-MolinaVenegas2017}{}%
Molina-Venegas, R., and M. A. Rodriguez. 2017. Revisiting phylogenetic
signal; strong or negligible impacts of polytomies and branch length
information? BMC evolutionary biology 17:53.

@@

\leavevmode\hypertarget{ref-Pavoine2012}{}%
Pavoine, S., and C. Ricotta. 2012. Testing for phylogenetic signal in
biological traits: The ubiquity of cross-product statistics. Evolution:
International Journal of Organic Evolution 67:828--840.

@@

\leavevmode\hypertarget{ref-Revell2012}{}%
Revell, L. J. 2012. Phytools: An r package for phylogenetic comparative
biology (and other things). Methods in Ecology and Evolution 3:217--223.
